Optimal strategies of influence in structured populations

### 2015-03-25 paper/manuscript.tex
correlation

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
  \frac{d \ln \tau}{d \sigma} = N \beta \sigma^{-2}
  \label{cost1}
\end{equation}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{figures/fig-chi.eps} 
%\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
%\label{fig1}
%\end{figure*}
We need to find $\tau$.
%\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

From Eqn.~\ref{lapdist}
\begin{align}
P\left(g_i>g_c \right) =
\frac{1}{2}\exp{\left[1/\sigma - \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }.%&\mbox{if } \\
%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%\end{cases}
\end{align}
(This expression is true if $x_L \le 1/2 - 1/(\sigma\ln\chi)$. As we are interested in these probabilities only when $x$ is small we restrict our attention to this scenario.) Given this equation we may write Eqn.~\ref{up} as
\begin{equation}
t^+(x) = (1-x)  e^{1/\sigma } \mathcal{F}(x)
\end{equation}
where 
\begin{equation}
 \mathcal{F}(x) = \frac{1}{2}\sum_{x_{L}}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}P\left(x_{L}|x\right)
\end{equation}
and similarly 
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^-(x) =x \left(1-  e^{1/\sigma } \mathcal{F}(x)\right)
\end{equation}
Taking the derivative of Eqn.~\ref{laplace_cost} and employing a Maclaurin series expansion leaves
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx N \sigma^{-2} \int_{x_1}^{x_2} \left(1 + \sum_{n=1}^{\infty} \left(e^{1/\sigma } \mathcal{F}(y)\right)^n  \right) dy
\label{laplace_cost2}
\end{equation}
The extrema of the potential, $x_1,x_2$, are at the points where $t^+(x)=t^-(x)$, i.e. where $x=e^{1/\sigma } \mathcal{F}(x)$. Since $e^{1/\sigma } \mathcal{F}(x)<x$ $\forall x \in [x_1,x_2]$ the summation terms in the integrand are all strictly less than $x$, hence
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx  N \sigma^{-2} \left({x_2}-{x_1}\right) + \mathcal{O}\left( \left(x_2-x_1\right)^2\right)
\label{laplace_cost2}
\end{equation}
Combining Eqns.~\ref{cost1} and~\ref{laplace_cost2} gives the location of minimal cost as
\begin{equation}
  \beta =  \left({x_2}-{x_1}\right) 
\label{laplace_cost_final}
\end{equation}
where in deriving this expression we have neglected terms of $\mathcal{O}(1/N)$ and $\mathcal{O}\left( \left(x_2-x_1\right)^2\right)$
Optimal strategy is therefore achieved by setting $\sigma$ so that the difference between the two extrema is equal to $\beta$.
%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.

>>>>

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
  \frac{d \ln \tau}{d \sigma} = N \beta \sigma^{-2}
  \label{cost_1}
\end{equation}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{figures/fig-chi.eps} 
%\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
%\label{fig1}
%\end{figure*}
We need to find $\tau$.
%\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

From Eqn.~\ref{lapdist}
\begin{align}
P\left(g_i>g_c \right) =
\frac{1}{2}\exp{\left[1/\sigma - \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }.%&\mbox{if } \\
%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%\end{cases}
\end{align}
(This expression is true if $x_L \le 1/2 - 1/(\sigma\ln\chi)$. As we are interested in these probabilities only when $x$ is small we restrict our attention to this scenario.) Given this equation we may write Eqn.~\ref{up} as
\begin{equation}
t^+(x) = (1-x)  e^{1/\sigma } \mathcal{F}(x)
\end{equation}
where 
\begin{equation}
 \mathcal{F}(x) = \frac{1}{2}\sum_{x_{L}}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}P\left(x_{L}|x\right)
 \label{social}
\end{equation}
and similarly 
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^-(x) =x \left(1-  e^{1/\sigma } \mathcal{F}(x)\right)
\end{equation}
Taking the derivative of Eqn.~\ref{laplace_cost} and employing a Maclaurin series expansion leaves
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx N \sigma^{-2} \int_{x_1}^{x_2} \left(1 + \sum_{n=1}^{\infty} \left(e^{1/\sigma } \mathcal{F}(y)\right)^n  \right) dy
\label{laplace_cost2}
\end{equation}
The extrema of the potential, $x_1,x_2$, are at the points where $t^+(x)=t^-(x)$, i.e. where $x=e^{1/\sigma } \mathcal{F}(x)$. Since $e^{1/\sigma } \mathcal{F}(x)<x$ $\forall x \in [x_1,x_2]$ the summation terms in the integrand are all strictly less than $x$, hence
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx  N \sigma^{-2} \left({x_2}-{x_1}\right) + \mathcal{O}\left( \left(x_2-x_1\right)^2\right)
\label{laplace_cost2}
\end{equation}
Combining Eqns.~\ref{cost_1} and~\ref{laplace_cost2} gives the location of minimal cost as
\begin{equation}
  \beta =  \left({x_2}-{x_1}\right) 
\label{laplace_cost_final}
\end{equation}
where in deriving this expression we have neglected terms of $\mathcal{O}(1/N)$ and $\mathcal{O}\left( \left(x_2-x_1\right)^2\right)$
Optimal strategy is therefore achieved by setting $\sigma$ so that the difference between the two extrema is equal to $\beta$.

Numerical calculations of $x_2-x_1$ and the optimal strategy are shown in Fig.~\ref{figx2x1}. More intuition into the underlying processes that control the distance between the inflexion points may be attained by performing an expansion of the social information, $\mathcal{F}(x)$, in terms of the moments of the sampling process. Taylor expanding the inner term %$\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}$ 
of Eqn.~\ref{social} around $x_L=x$ leads to % allows us to replace the summation over all values of $x_L$ with its mean and variance, i.e.
\begin{equation}
\mathcal{F}(x) = \frac{1}{2}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x\right)\right]} \left( 1 + \frac{ \ln(\chi)^2}{2}\mbox{Var}(x_L) \right)
\end{equation}
Effectively this equation approximates the process of observing a restricted subset of the population. The observations of opinions of neighbors represents an estimate, $x_L$, of the true population average opinion, $x$. This random process of observations may be approximated in terms of the mean and variance of the sampling. While the underlying network structure is complex and non-Markovian, it may be modelled as correlated binomial, in essence we model the process as sampling with pairwise correlation, when there is transitivity near-neighbors are likely to have copied one another hence opinions will be clustered. Following ~\cite{witt2004moody}, the variance of a correlated binomial is
\begin{equation}
\mbox{Var}(x_L) = x(1-x)\left( \rho + \frac{1-\rho}{N_s} \right)
\end{equation}
where $\rho$ is the pairwise correlation coefficient. Assuming the network structure and conformity create a positive correlation of opinions amongst neighbors, then we may define an increased likelihood that neighbor $j$ is in state $S_j=1$, given $S_k=1$, so that
\begin{equation}
P(S_j=1|S_k=1) = x(1+\alpha)
\end{equation}
where $\alpha$ is determined by the transitivity (probability $k$ is in the neighborhood of $j$) and conformity. This means 
\begin{equation}
\rho = \frac{x\alpha}{1-x}
\end{equation}
and
\begin{align}
\mathcal{F}(x) &= \frac{1}{2}\exp{\left[ -  \ln\chi  \left(\frac{1}{2}-x\right)\right]} \times \nonumber \\ & \left( 1 + \frac{ \ln(\chi)^2}{2}\left(\frac{x(1-x(1+\alpha))}{N_s} +  \alpha x^2  \right)   \right).
\end{align}
Solutions to $\mathcal{F}(x)=x$ may be constructed with a perturbation method, with small parameters $\alpha$ and $1/N_s$. We begin with 
\begin{equation}
x_1 \sim x^0_1 + \alpha x^1_1 + 1/N_s x^2_1
\end{equation}
The zeroth order solution is for the mean field where there is no clustering and no noise in the sampling process, hence
\begin{align}
x^0  = \frac{1}{2}\exp{\left[ 1/\sigma -  \ln\chi  \left(\frac{1}{2}-x^0\right)\right]} 
\end{align}
For small x
\begin{equation}
x^0_1 = \frac{1}{\exp{\left[ 1/\sigma -  \ln\chi  \left(\frac{1}{2}\right)\right]}  - \ln\chi}
\end{equation}
%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.
